---
title: "Alerts Analysis"
author: "Ricardo Maceiras"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This dataset shows a simulated sample of alerts generated by specific scenarios will be analyzed, these scenarios were simulated for two types of customers, Private Banking and LC&FI.

The topics which were analyzed:

* Alert Volume and Types
* Alert Resolution Times
* Customer Risk Categories
* PEP Status
* Industry alerts
* Additional Analysis



```{r}
library(tidyverse)
library(readxl)

Alerts_Dataset <- read_excel("alerts.xlsx")
Alerts_Dataset

```

## 1) Alert Volume and Types

### 1.a Alert volume

```{r}
alert_typesVOL <- Alerts_Dataset %>% 
  select(AlertType) %>% 
  group_by(AlertType) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100,2)) %>% 
  arrange(-Total_values)

alert_typesVOL


colors <- c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17BECF", "#FFC0CB", "#FF1493", "#1F497D", "#00BFFF")

alert_typesVOL_chart <- ggplot(alert_typesVOL, aes(x = "", y = Percentage, fill = AlertType)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(x = NULL, y = NULL, title = "Alert Type Distribution") +
  scale_fill_manual(values = colors)

alert_typesVOL_chart
```

According to the previous table and chart:

* The alert type "Unusual behavior" has the highest volume of alerts in the sample, approximately 25.2% of the total with 2563 alerts.

* The second and third highest type of alerts are "Credit Cards" and "New Destinations with high turnover" with 1997and 1543 alerts respectively, accounting 19,6% and 15.2% of the total.

* Apparently, the alert types with the lowest numbers of alerts are "Unusual Cash Behaviour" and "International Transfers"
with 64 and 34 alerts respectively, accounting 0.63% and 0.33% of the total.


### 1.b Alert volume: Private Banking vs. LC&FI Analysis

```{r}
Type_status <- Alerts_Dataset %>% 
  select(Type, AlertType) %>% 
  group_by(Type, AlertType) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100,2))

Type_status


Type_status_arrange <- Alerts_Dataset %>% 
  select(Type, AlertType) %>% 
  group_by(Type, AlertType) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100,2)) %>% 
  arrange(-Total_values)

Type_status_arrange
```

According to the previous records:

* The highest types of alerts when the customer is Private Banking are "Unusual behaviour", "New Destinations with high turnover", " Existing Accounts" with 1715, 1543 and 220 alerts respectively, accounting 44.7%, 40.2% abd 5.73% of the total within the group.

* The highest types of alerts when the customer is LC&FI are "Credit Cards", "Check Countries List", "Unusual behaviour" with 1997, 1382, 848 alerts, accounting 31.5%, 21.8%, 13.4% of the total within the group.

* "Credit Cards" alerts occur when the customer is LC&FI.
* "Unusual behavior" alerts are the type with most appearances in both Private Banking and LC&FI.



## 2) Alert Resolution Times

The elapsed time from the date when 2nd line started the investigation to the Date when 2d line closed the investigation
will be analyzed


```{r}
Alerts_Dataset$CaseOpen <- as.POSIXct(Alerts_Dataset$CaseOpen, format="%Y-%m-%d %H:%M:%S")
Alerts_Dataset$CaseClosed <- as.POSIXct(Alerts_Dataset$CaseClosed, format="%Y-%m-%d %H:%M:%S")
Alerts_Dataset$CaseReported <- as.POSIXct(Alerts_Dataset$CaseReported, format="%Y-%m-%d %H:%M:%S")

```

### 2.a Alert Resolution Times: Case open - Case closed

```{r}

alert_times_CASOP_CASCL <- Alerts_Dataset %>% 
  select(AlertType, CaseOpen, CaseClosed, Type) %>% 
  mutate(Resolution_hours = round(difftime(CaseClosed, CaseOpen, units = "hours"),2)) %>% 
  drop_na()

alert_times_CASOP_CASCL_Summary <- alert_times_CASOP_CASCL %>% 
  select(AlertType,Resolution_hours) %>% 
  group_by(AlertType) %>% 
  summarize(Average_resolution_hours= round(mean(Resolution_hours),2)) %>% 
  arrange(-Average_resolution_hours)

alert_times_CASOP_CASCL_Summary

max_value <- as.numeric(max(alert_times_CASOP_CASCL_Summary$Average_resolution_hours)) + 1000

alert_times_CASOP_CASCL_Summary_chart <- ggplot(alert_times_CASOP_CASCL_Summary, aes(x = AlertType, y = Average_resolution_hours)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Alert Type", y = "Average Resolution Hours", title = "Average Resolution Hours by Alert Type") +
  scale_y_continuous(limits = c(0, max_value), breaks = seq(0, max_value, by = 1000))

alert_times_CASOP_CASCL_Summary_chart 

```

According to the previous table and chart:

* Apparently, the alert types which investigation takes the most time to be closed when 2nd line start it are "Credit Cards" alerts which might take  12887.43 hour on average and "Cash" alerts with  6799.32 hours on average.

* Otherwise, the alert types which investigation takes the lowest period of time to be closed by 2nd libe are "Unusual behaviour" which might take 0.17 hours on average and Awakening Account with 0.03 hours on average.

* "Unusual behaviour" and "Credit Cards" are the type with the largest number of alerts, being "Unusual behaviour" cases resolved quickly and efficiently and "Credit Cards" cases not as efficient as others.


### 2.a.a Alert Resolution Times: Private Banking vs. LC&FI Analysis

```{r}
Types_CASOP_CASCL_Summary <- alert_times_CASOP_CASCL %>% 
  select(Type, AlertType,Resolution_hours) %>% 
  group_by(Type, AlertType) %>% 
  summarize(Average_resolution_hours= round(mean(Resolution_hours),2)) 

Types_CASOP_CASCL_Summary

```

According to the table:

* Usually the cases of Private Banking are resolved faster and more efficiently than cases of LC&FI by the second line once opened the investigation.


### 2.b Alert Resolution Times: Case open - Case Reported

```{r}
alert_times_CASOP_CASRE <- Alerts_Dataset %>% 
  select(AlertType, CaseOpen, CaseReported,Type) %>% 
  mutate(Resolution_hours = round(difftime(CaseReported, CaseOpen, units = "hours"),2)) %>% 
  drop_na()
  
alert_times_CASOP_CASRE_Summary <- alert_times_CASOP_CASRE %>% 
  select(AlertType,Resolution_hours) %>% 
  group_by(AlertType) %>% 
  summarize(Average_resolution_hours= round(mean(Resolution_hours),2)) %>% 
  arrange(-Average_resolution_hours)

alert_times_CASOP_CASRE_Summary

max_value2 <- as.numeric(max(alert_times_CASOP_CASRE_Summary$Average_resolution_hours)) + 10

alert_times_CASOP_CASRE_Summary_chart <- ggplot(alert_times_CASOP_CASRE_Summary, aes(x = AlertType, y = Average_resolution_hours)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Alert Type", y = "Average Resolution Hours", title = "Average Resolution Hours by Alert Type") +
  scale_y_continuous(limits = c(0, max_value2), breaks = seq(0, max_value2, by = 10))

alert_times_CASOP_CASRE_Summary_chart

Types_CASOP_CASRE_Summary <- alert_times_CASOP_CASRE %>% 
  select(Type, AlertType,Resolution_hours) %>% 
  group_by(Type, AlertType) %>% 
  summarize(Average_resolution_hours= round(mean(Resolution_hours),2)) 

Types_CASOP_CASRE_Summary

```

According to the previous table and chart:

* "Listed High Risk Banks" and "Cash" are the alert types that take the most time to be sent to FIU with 172.42 and 125.81 hours on average respectively. Nonetheless, "Unusual behaviour" alerts are sent in 0.05 hours on average.

* Most of the alerts reported to FIU occurred when the customer is LC&FI. Moreover, LC&FI alerts take much longer hours than Private Banking alerts.

### 2.c Alert Resolution Times: First line to Second line

```{r}
sum_na= sum(is.na(Alerts_Dataset$CaseOpen)) 

alert_times_FIRST_SECOND <- Alerts_Dataset %>% 
  group_by(AlertType) %>% 
  summarize(Percentage_alerts = round(mean(is.na(CaseOpen)) * 100,2)) %>% 
  arrange(desc(Percentage_alerts))

alert_times_FIRST_SECOND

```

* The vast majority of the alerts pass from first line to be investigated by second line.

* 94.2 % of "Credit Cards" alerts pass to second line, being the lowest ratio among alert types.


## 3) Customer Risk Categories

```{r}
Risk_categories <- Alerts_Dataset %>% 
  select(CusRiskCategory) %>% 
  group_by(CusRiskCategory) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100,2)) %>% 
  arrange(-Total_values)

Risk_categories

```

* 0,53% of the total are NULL values, therefore, can be omitted

### 3.a Customer Risk Categories: Summary of risk categories and alert type

```{r}
Risk_categories_AlertyType <- Alerts_Dataset %>% 
  select(AlertType, CusRiskCategory) %>% 
  filter(CusRiskCategory != "NULL") %>% 
  group_by(AlertType, CusRiskCategory) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100, 2))

Risk_categories_AlertyType

colors <- c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", "#8C564B", "#E377C2", "#7F7F7F", "#BCBD22", "#17BECF", "#FFC0CB", "#FF1493", "#1F497D", "#00BFFF")

Risk_categories_AlertyType_chart <- ggplot(Risk_categories_AlertyType, aes(x = Percentage, y = AlertType, fill = CusRiskCategory)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Percentage", y = "AlertType", title = "Percentage of Total Values by Alert Type and Customer Risk Category") +
  scale_fill_manual(values = colors)

Risk_categories_AlertyType_chart
```

According to the previous chart and table:

* Most of the alert types are categorized as "Medium Risk" with the exception of "Unusual cash behaviour" which most of their alerts are "Lower Risk" and "Pep monitoring" where most of their alerts are not specified, therefore, hindering the sample.



### 3.a Customer Risk Categories: Summary of risk categories and type company

```{r}
Risk_categories_Type <- Alerts_Dataset %>% 
  select(Type, CusRiskCategory) %>% 
  filter(CusRiskCategory != "NULL") %>% 
  group_by(Type, CusRiskCategory) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100, 2))

Risk_categories_Type

Risk_Categories_lcfi_chart <- ggplot(Risk_categories_Type%>% filter(Type == "lcfi"), aes(x = "", y = Percentage, fill = CusRiskCategory)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(x = NULL, y = NULL, title = "Customer Risk Distribution for Type lcfi") +
  scale_fill_manual(values = colors)+
  geom_label(aes(label = paste0(Percentage, "%")), position = position_stack(vjust = 0.5))

Risk_Categories_lcfi_chart

Risk_Categories_pb_chart <- ggplot(Risk_categories_Type%>% filter(Type == "pb"), aes(x = "", y = Percentage, fill = CusRiskCategory)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(x = NULL, y = NULL, title = "Customer Risk Distribution for Type pb") +
  scale_fill_manual(values = colors)+
  geom_label(aes(label = paste0(Percentage, "%")), position = position_stack(vjust = 0.5))

Risk_Categories_pb_chart

```

* Most of the alerts of Private Banking and LC&FI are categorized as "Medium Risk".

* Private Banking has more "Lower Risk" alerts and LC&FI has more "Higher Risk" alerts.


## 4) PEP Status

```{r}
Pep_status <- Alerts_Dataset %>% 
  select(PEP) %>% 
  group_by(PEP) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100,2)) %>% 
  arrange(-Total_values)
 
Pep_status  

Pep_status_AlertyType <- Alerts_Dataset %>% 
  select(AlertType, PEP) %>% 
  mutate(PEP = ifelse(is.na(PEP), "Missing Value", PEP)) %>% 
  group_by(AlertType, PEP) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100, 2))

Pep_status_AlertyType


Pep_status_chart <- ggplot(Pep_status, aes(x = "", y = Percentage, fill = PEP)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(x = NULL, y = NULL, title = "PEP Distribution") +
  scale_fill_manual(values = colors)+
  geom_label(aes(label = paste0(Percentage, "%")), position = position_stack(vjust = 0.5))

Pep_status_chart

Pep_status_AlertyType_chart <- ggplot(Pep_status_AlertyType, aes(x = Percentage, y = AlertType, fill = PEP)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Percentage", y = "Alert Type", title = "Percentage of Total Values by Alert Type PEP") +
  scale_fill_manual(values = colors)

Pep_status_AlertyType_chart

```

According to the charts and tables:

* There is a problem of biased data, 62.27 of the values are NA and 9.22% are NULL, which difficult the analyzis, NA values cannot be removed, otherwisde the results would not be representative within the sample.

* However, "New Destinations with high turnover", "Recurring In-Out scenario","International Transfers" alert types can be analyzed, being the majority of their alerts, negative PEP.


## 5) Industry alerts

```{r}

Industry_code <-Alerts_Dataset %>% 
  select(IndustryCode) %>% 
  filter(IndustryCode != "NULL" & IndustryCode != 0) %>% 
  group_by(IndustryCode) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100, 2)) %>% 
  arrange(-Total_values)

## adding other dataset
additional_info <- read_excel("additional_info.xlsx", sheet = 2)

Additional_info_edit <- additional_info %>% 
  rename(IndustryCode = "Industry Code",RiskScore= "Risk Score")


Industry_risk_leftjoin <- left_join(Industry_code, Additional_info_edit, by = "IndustryCode")
Industry_risk_leftjoin

Industry_risk <- inner_join(Industry_code, Additional_info_edit, by = "IndustryCode")

Industry_risk_General <- inner_join(Alerts_Dataset, Additional_info_edit, by = "IndustryCode")

Industry_risk_General

Risk_segment <- Industry_risk_General %>% 
  select(Type, Segment) %>% 
  group_by(Type, Segment) %>% 
  summarize(Total_values = n()) %>% 
  mutate(Percentage = round(Total_values / sum(Total_values) * 100, 2))

Risk_segment


Risk_segment_chart <- ggplot(Risk_segment%>% filter(Type == "lcfi"), aes(x = "", y = Percentage, fill = Segment)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(x = NULL, y = NULL, title = "Risk segment distribution for Type lcfi") +
  scale_fill_manual(values = colors)+
  geom_label(aes(label = paste0(Percentage, "%")), position = position_stack(vjust = 0.5))

Risk_segment_chart


```

Observations:

* The majority of the alerts are not categorized by industry code and segment. Moreover, he metadata does not contain all industry codes of the data set.

* There were only 80 values that matched with the Industry code, being all of them LC&FI with "High Risk"



## 6) Additional Analysis

What analysis you think should be included in the qualitative validation part of Transaction Monitoring model? What areas of TM models are essential to analyse?.


Analysis to be included:

* PEP details: add information on the role or position held by a PEP so that we can understand the risk associated to this person in concern.

* PEP exposure in High risk countries: Determine whether the PEP has connections or businesses in regions associated to illicit activities

* Regional Risk Assessment: Analyze each industry and where their operations are held in as well as the exposure to high-risk jurisdictions.


Essential areas:

I consider important the analysis of PEP as well as Industry risk. Nonetheless, during the analysis i found these observations:

* Most of the values in this PEP model are missing values or NULL, instead of Non PEP/PEP. There should be a better mapping.

* Most of the industry codes are missing values and the meta data (additional information) does not map all the codes, therefore, it biases the sample.




